Mining Scenario/Transaction Pairs in HW/SW Co-verification Using Symbolic Execution

### 2017-04-04 paper.tex
section 3.2

%This interaction pattern exhibits producer-consumer 
%relationship, which is also evident in case of our UART example in 
%Figure~\ref{fig:interleaving1}.  
%
A communicating pair of executions of 
$\mathcal{FT}$ and $\mathcal{HT}$ is an {\em interaction}, which 
takes place through shared variables (or shared registers)
only.  An interaction is said to exhibit {\em producer-consumer}
interaction pattern if 1) the set of variables $V$ written by 
the execution of $\mathcal{FT}$ and and the set of variables 
$V'$ written by the execution of $\mathcal{HT}$ are disjoint 
($V \cap V' = \emptyset$), (2) and the executions 
of $\mathcal{FT}$ and $\mathcal{HT}$ consumes values 
in $V'$ and $V$, respectively.
%
Figure~\ref{fig:interleaving1} illustrates the producer-consumer relationship 
in UART IP. The FW CFG is shown in the left and the 
HW RTL code snippet is shown in the right. We abuse notations and 
represent $\mathcal{HT}$ in the form of RTL and $\mathcal{FT}$ as CFG
for our illustration purpose. Here, $\mathcal{HT}$ is the producer of interrupt 
status signal, $irq\_o$, which is read by $\mathcal{FT}$ in state $s_1$. 
The signal $irq\_o$ is  triggered when transmitter buffer is 
{\em EMPTY} or receiver buffer is {\em FULL}. 
$\mathcal{HT}$ also produces interrupt status by writing to the 
memory mapped register, $UART\_IS$, which is consumed by 
$\mathcal{FT}$ in state $s_2$.  In another interaction, $\mathcal{HT}$
sends a data-packet to $UART\_TR$ which is consumed by $\mathcal{FT}$ in state $s_7$.
Similarly, $\mathcal{FT}$ is the producer of $data$, $(outb(data,UART\_TR))$, 
in state $s_6$ which is consumed by $\mathcal{HT}$.  
%
\begin{figure}[t]
\scalebox{.55}{\import{figures/}{interleaving.pspdftex}}
\caption{Producer Consumer Interactions Pattern
\label{fig:interleaving1}}
\end{figure}
%
\begin{table}
\begin{center}
{\scriptsize
\begin{tabular}{|c|c|c|}
\hline
  Firmware & Hardware  \\
  Transaction ($\mathcal{ST}$) & Transaction ($\mathcal{HT}$) \\ 
\cline{1-2}
   $s_1 \xrightarrow[]{irq_o} s_2 \xrightarrow[\mathcal{C}]{inb()} s_4
   \xrightarrow[]{status == 0x0c} s_6 \xrightarrow[\mathcal{P}]{outb()} s_9 \rightarrow
   s_{11} \rightarrow s_{12}$  & $s_2 \xrightarrow[\mathcal{C}]{inb()} s_4\;\; uart\_{rx}$
   \\ 
   & $s_6 \xrightarrow[]{outb()} s_9\;\; uart\_{tx}$ \\  \hline

@@

   $s_1 \xrightarrow[]{\neg irq_o} s_3 \xrightarrow[]{wb\_{idle()}}
   s_5 \xrightarrow[]{wb\_{idle()}} s_8 \rightarrow
   s_{12}$ & $s_3 \xrightarrow[]{wb\_{idle()}} s_5\;\; uart\_{top}$ \\ 
   & $s_5 \xrightarrow[]{wb\_{idle()}} s_8\;\; uart\_{top}$ \\  \hline
\end{tabular}
}
\end{center}
\caption{\scriptsize Firmware Transactions and its interacting HW transaction}
\label{table:stats}
\end{table}  
%
This interaction behavior can be analysed 
to produce a sequential single-threaded co-verification model.

\textbf{Step 3: Co-verification Model}
An unified co-verification model is constructed by sequentializing 
HW and its interacting FW threads, thereby generating a sequential 
single-threaded co-verification model in C.  
%
We now formally define a single-threaded sequential co-verification model, 
denoted by $\mathcal{M}_{seq}$, obtained by sequentializing the HW and FW 
models.
%
\[ 
\begin{array}[t]{@{}lll}
  \mathcal{M}_{seq} & {:}{:}{=} & \mathit{Procedure} \\
\mathit{Procedure} & {:}{:}{=} & \mathit{Statement} \mid \mathit{Procedure} \\
\mathit{Statement} & {:}{:}{=} & CStatement \mid \mathit{moduleName(Var_1,\dots,Var_n)} \\
\mathit{CStatement} & {:}{:}{=} & x {:}{=} exp \mid \mathit{ITE}(b, s_1,s_2) \mid \mathit{s_1;s_2} \mid loop\{s\} \\ 
\end{array}
\]
%
Consider sets of expressions $Exp$ and Boolean expressions $BExp$ 
over variables $Var$ of $\mathcal{M}_{seq}$ such that $(exp \in Exp)$ and 
$(b \in BExp)$. A $\mathit{CStatement}$ is an assignment,
conditional, sequential concatenation or a loop.
$\mathit{moduleName(Var_1,\dots,Var_n)}$ invokes the top level 
procedure of the $\mathcal{SN}$. Recall that $\mathcal{SN}$ is a
software netlist representation of HW RTL models.
The interface variables $\{Var\_1,\dots,Var\_n\}$ 
(input, output and inout ports) are declared inside the 
FW to communicate with $\mathcal{SN}$. \\ 
%

>>>>

%This interaction pattern exhibits producer-consumer 
%relationship, which is also evident in case of our UART example in 
%Figure~\ref{fig:interleaving1}.  
%
A communicating pair of executions of 
$\mathcal{FT}$ and $\mathcal{HT}$ is an {\em interaction}, which 
takes place through shared variables (or shared registers)
only.  An interaction is said to exhibit {\em producer-consumer}
interaction pattern if 1) the set of variables $V$ written by 
the execution of $\mathcal{FT}$ and and the set of variables 
$V'$ written by the execution of $\mathcal{HT}$ are disjoint 
($V \cap V' = \emptyset$), (2) and the executions 
of $\mathcal{FT}$ and $\mathcal{HT}$ consumes values 
in $V'$ and $V$, respectively.
%
Figure~\ref{fig:interleaving1} illustrates the producer-consumer relationship 
in UART IP. The FW CFG is shown in the left and the 
HW RTL code snippet is shown in the right. We abuse notations and 
represent $\mathcal{HT}$ in the form of RTL and $\mathcal{FT}$ as CFG
for our illustration purpose. Here, $\mathcal{HT}$ is the producer of interrupt 
status signal, $irq\_o$, which is read by $\mathcal{FT}$ in state $s_1$. 
The signal $irq\_o$ is  triggered when transmitter buffer is 
{\em EMPTY} or receiver buffer is {\em FULL}. 
$\mathcal{HT}$ also produces interrupt status by writing to the 
memory mapped register, $UART\_IS$, which is consumed by 
$\mathcal{FT}$ in state $s_2$.  In another interaction, $\mathcal{HT}$
sends a data-packet to $UART\_TR$ which is consumed by $\mathcal{FT}$ in state $s_7$.
Similarly, $\mathcal{FT}$ is the producer of $data$, $(outb(data,UART\_TR))$, 
in state $s_6$ which is consumed by $\mathcal{HT}$.  
%
\begin{figure}[t]
\scalebox{.55}{\import{figures/}{interleaving.pspdftex}}
\caption{Producer Consumer Interactions Pattern
\label{fig:interleaving1}}
\end{figure}
%
Table~\ref{transaction} shows the FW transactions $(\mathcal{FT})$ and its 
interacting HW transactions $(\mathcal{HT})$ from Figure~\ref{fig:interleaving1}. 
The description above the arrows 
in column 1 shows the events in $\mathcal{FT}$ and the letter below the
arrows denotes the event type, producer $(\mathcal{P})$ or consumer 
$(\mathcal{C})$. Arrows with empty description below denotes reading of interface
variables or local variables in $\mathcal{FT}$.  Arrows with no description
simply denotes control-flow with no events.  The entires in the right column of
Table~\ref{transaction} shows the corresponding transaction in HW modules. We
simply denote a transaction in HW module with the name of the module,
transmitter module in UART ($uart\_tx$), receiver module in UART ($uart\_rx$), or
the top UART module ($uart\_top$).
%
\begin{table}
\begin{center}
{\scriptsize
\begin{tabular}{|c|c|c|}
\hline
  Firmware & Hardware  \\
  Transaction ($\mathcal{ST}$) & Transaction ($\mathcal{HT}$) \\ 
\cline{1-2}
   $s_1 \xrightarrow[]{irq_o} s_2 \xrightarrow[\mathcal{C}]{inb()} s_4
   \xrightarrow[]{status == 0x0c} s_6 \xrightarrow[\mathcal{P}]{outb()} s_9 \rightarrow
   s_{11} \rightarrow s_{12}$  & $s_2 \xrightarrow[\mathcal{C}]{inb()} s_4\;\; uart\_{rx}$
   \\ 
   & $s_6 \xrightarrow[]{outb()} s_9\;\; uart\_{tx}$ \\  \hline

@@

   $s_1 \xrightarrow[]{\neg irq_o} s_3 \xrightarrow[]{wb\_{idle()}}
   s_5 \xrightarrow[]{wb\_{idle()}} s_8 \rightarrow
   s_{12}$ & $s_3 \xrightarrow[]{wb\_{idle()}} s_5\;\; uart\_{top}$ \\ 
   & $s_5 \xrightarrow[]{wb\_{idle()}} s_8\;\; uart\_{top}$ \\  \hline
\end{tabular}
}
\end{center}
\caption{\scriptsize Firmware Transactions and its interacting HW transaction}
\label{transaction}
\end{table}  
%

\textbf{Step 3: Co-verification Model}
An unified co-verification model is constructed by instrumenting  
the software netlist model of HW RTL inside the FW thread, thereby 
generating a sequential single-threaded co-verification model in C.  
%
We now formally define a our co-verification model, denoted by 
$\mathcal{M}_{seq}$.\rmcmt{refer moduleName} as software netlist.
%
\[ 
\begin{array}[t]{@{}lll}
  \mathcal{M}_{seq} & {:}{:}{=} & \mathit{Procedure} \\
\mathit{Procedure} & {:}{:}{=} & \mathit{Statement} \mid \mathit{Procedure} \\
\mathit{Statement} & {:}{:}{=} & CStatement \mid \mathit{moduleName(Var_1,\dots,Var_n)} \\
\mathit{CStatement} & {:}{:}{=} & x {:}{=} exp \mid \mathit{ITE}(b, s_1,s_2) \mid \mathit{s_1;s_2} \mid loop\{s\} \\ 
\end{array}
\]
%
Consider sets of expressions $Exp$ and Boolean expressions $BExp$ 
over variables $Var$ of $\mathcal{M}_{seq}$ such that $(exp \in Exp)$ and 
$(b \in BExp)$. A $\mathit{CStatement}$ is an assignment,
conditional, sequential concatenation or a loop.
$\mathit{moduleName(Var_1,\dots,Var_n)}$ invokes the top level 
procedure of the $\mathcal{SN}$. Recall that $\mathcal{SN}$ is a
software netlist representation of HW RTL models.
The interface variables $\{Var\_1,\dots,Var\_n\}$ 
(input, output and inout ports) are declared inside the 
FW to communicate with $\mathcal{SN}$. \\ 
%
